\documentclass{masterthesis}

\usepackage[
    pdftex,
    pdftitle={Learning relational models from networked data on the web},
    pdflang={en}
]{hyperref}
\usepackage{cmap}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{dirtytalk}
\usepackage{subfigure}
\usepackage{bbold}
\usepackage[english]{babel}

\DeclareMathOperator*{\E}{\mathbb{E}}
\setlist[itemize]{topsep=0pt}
\setlist[enumerate]{topsep=0pt, itemsep=0pt}

\begin{document}

\title{Learning relational models \\ from networked data on the web}

\author{Davide Riva}

\advisor{Annalisa Barla, Davide Garbarino}

\examiner{Nicoletta Noceti}

\maketitle

\begin{abstract}
    Network inference is the process of uncovering relations between data, 
    particularly useful in biology to reconstruct Gene Regulatory Networks (GRN). 
    Supervised methods were also proposed in the past 
    to summarize a network of documents and predict links between them.

    In this thesis, we propose an unsupervised approach 
    to extract an undirected and unweighted graph given a collection of texts.
    In particular, we present a pipeline to construct 
    and use a dataset composed of pages of a given website.
\end{abstract}

\dedication{
    I wish to express my deepest gratitude to all students and professors
    who went along with me on this journey.
    Among those, a special thanks to Eros Viola,
    Riccardo Bianchini, Olga Matthiopoulou, Marco Rando,
    Andrea Canepa, Larbi Touijer, Federico Lodovici,
    Roberto Castellotti, Antonio Vivace,
    Marco Orsingher, Michele Pugno, 
    Adrian Castro, 
    Annalisa Barla, Nicoleta Noceti and Davide Garbarino.

    In addition, I would like to thank Jordan Boyd-Graber and
    David M. Blei who released for free their lessons, talks and material
    which were useful during the writing of this thesis.

    I am also indebted to Fabio Antonio Stella,
    which made me fall in love with this subject
    and played an important role in the choices I made that led me here.

    I wish to acknowledge the support and great love of Laura
    and her family, which helped me tackle this degree course which is going to end.

    Last but not least,
    I would like to recognize the invaluable economic assistance
    that my family provided during my study.

    Thank you all.
}

\setcounter{tocdepth}{1}
\tableofcontents

\input{chapters/introduction.tex}
\input{chapters/wordEmbedding.tex}
\input{chapters/topicModeling.tex}
\input{chapters/networks.tex}
\input{chapters/experiments.tex}
\input{chapters/conc_and_future.tex}

\appendix
\input{chapters/appendix/nn.tex}
\input{chapters/appendix/probabilityDistributions.tex}
\input{chapters/appendix/statisticalDistances.tex}
\input{chapters/appendix/vi.tex}

\bibliographystyle{alphaurl}
\bibliography{bib}

\include{chapters/license}

\end{document}